Parametricity for Primitive Nested Types

### 2021-01-11 free-theorems-nesteds/20-mscs.tex
updated LMCS paper

We can prove by simultaneous induction that our interpretations of
types interact well with demotion of functorial variables to
non-functorial ones, along with other useful identities. Indeed, if
$\rho, \rho' : \setenv$, \,$f : \rho \to \rho'$, \,$\rho \phi = \rho
\psi = \rho' \phi = \rho' \psi$, \, $f \phi = f \psi = \id_{\rho
  \phi}$,\, $\Gamma; \Phi, \phi^k \vdash F$,\,
$\Gamma;\Phi,\ol{\alpha} \vdash G$,\, $\Gamma;\Phi,\alpha_1...\alpha_k
\vdash H$, and $\ol{\Gamma;\Phi \vdash K}$, then
\begin{gather}
\label{thm:demotion-objects}
\setsem{\Gamma; \Phi, \phi \vdash F} \rho = \setsem{\Gamma, \psi; \Phi
  \vdash F[\phi :== \psi] } \rho\\
\label{thm:demotion-morphisms}
\setsem{\Gamma; \Phi, \phi \vdash F} f = \setsem{\Gamma, \psi; \Phi
  \vdash F[\phi :== \psi]} f\\
\label{eq:subs-var}
\setsem{\Gamma;\Phi \vdash G[\ol{\alpha := K}]}\rho =
\setsem{\Gamma;\Phi,\ol{\alpha} \vdash G}\rho[\ol{\alpha := 
\setsem{\Gamma;\Phi \vdash K}\rho}]\\
\label{eq:subs-var-morph}
\setsem{\Gamma;\Phi \vdash G[\ol{\alpha := K}]}f =
\setsem{\Gamma;\Phi,\ol{\alpha} \vdash G}f[\ol{\alpha :=
\setsem{\Gamma;\Phi \vdash K}f}]\\
\label{eq:subs-const}
\setsem{\Gamma; \Phi \vdash F[\phi := H]}\rho
= \setsem{\Gamma; \Phi, \phi \vdash F}\rho
[\phi := \lambda \ol{A}.\, \setsem{\Gamma;\Phi,\overline{\alpha}\vdash
    H}\rho[\overline{\alpha := A}]] \\ 
\label{eq:subs-const-morph}
\setsem{\Gamma; \Phi \vdash F[\phi := H]}f
= \setsem{\Gamma; \Phi, \phi \vdash F}f
[\phi := \lambda \ol{A}.\,\setsem{\Gamma;\Phi,\overline{\alpha}\vdash
    H}f[\overline{\alpha := \id_{\ol{A}}}]] 
\end{gather}
Identities analogous to (\ref{thm:demotion-objects}) through
(\ref{eq:subs-const-morph}) hold for relational interpretations as well.

@@

\begin{figure*}
\hspace*{-3.1in}
\resizebox{0.46\linewidth}{!}{
\begin{minipage}[t]{0.5\textwidth}
\[\begin{array}{lll}
\dsem{\Gamma;\Phi \,|\, \Delta,x :F \vdash x : F} \rho& = &
\pi_{|\Delta|+1}\\
\dsem{\Gamma;\emptyset \,|\, \Delta \vdash L_{\overline \alpha} x.t :
  \Nat^{\overline 
    \alpha} \,F \,G}\rho & = &  \curry (\dsem{\Gamma;\overline \alpha
  \,|\, \Delta, x : F \vdash t: G}\rho[\overline{\alpha := \_}])\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t_{\overline K} s:
  G [\overline{\alpha := K}]}\rho & = & \eval \circ \langle
  \lambda d.\,(\dsem{\Gamma;\emptyset \,|\, \Delta \vdash t :
  \Nat^{\overline{\alpha}} \,F \,G}\rho\; d)_{\overline{\dsem{\Gamma;\Phi
      \vdash K}\rho}},\\ 
 & & \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\,
    \Delta \vdash s: F [\overline{\alpha := K}]}\rho \rangle\\ 
& & \\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \bot_F t : F} \rho& = &
!^0_{\dsem{\Gamma;\Phi \vdash F}\rho} \circ
  \dsem{\Gamma;\Phi~|~\Delta \vdash t : \zerot}\rho, \mbox{ where } \\
 & & \hspace*{0.1in} !^0_{\dsem{\Gamma;\Phi \vdash F}\rho}
\mbox{ is the unique morphism from } 0\\
 & & \hspace*{0.1in} \mbox{ to } \dsem{\Gamma;\Phi \vdash F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \top : \onet}\rho & = &
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1, \mbox{ where }
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1\\ 
& & \hspace*{0.1in} \mbox{ is the unique morphism from }
\dsem{\Gamma;\Phi\vdash \Delta}\rho \mbox{ to } 1\\ 
\dsem{\Gamma;\Phi \,|\, \Delta \vdash (s,t) : F \times G} \rho& = &
\dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F} \rho\times
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G} \rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_1 t : F} \rho& = &
\pi_1 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times G}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_2 t : G}\rho & = &
\pi_2 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times
  G} \rho\\
\dsem{\Gamma;\Phi~|~\Delta \vdash \cse{t}{x \mapsto l}{y \mapsto r} :
  K}\rho & = & \eval \circ \langle \curry \,[\dsem{\Gamma;\Phi
    \,|\, \Delta, x : F \vdash l : K}\rho,\\
   & & \hspace*{0.79in} \dsem{\Gamma;\Phi \,|\, \Delta, y
    : G \vdash r : K}\rho],\\
   & &  \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\, \Delta \vdash t :
  F + G} \rho\rangle\\   
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inl \,s: F + G} \rho& = &
\inl \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inr \,t: F + G}\rho & = & 
\inr \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G}\rho\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \map^{\ol{F},\ol{G}}_H
  : \Nat^\emptyset (\Nat^{\ol\beta,\ol\gamma} F\,G)
& = & \lambda d\, \ol\eta\,\ol{C}.\,
\dsem{\Gamma; \ol{\phi},\ol{\gamma}\vdash H}\id_{\rho[\ol{\gamma:=
      C}]}[\ol{\phi := \lambda \ol B. \eta_{\ol B\, \ol C}}]\\ 
\hspace*{0.5in}
  (\Nat^{\ol{\gamma}}\,H[\ol{\phi :=_{\ol{\beta}} F}]\,H[\ol{\phi
      :=_{\ol{\beta}} G}]) \rrbracket^{\mathsf D} \rho & &\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
Nat^{\ol{\beta}} \, H[\phi := (\mu \phi.\lambda {\overline
    \alpha}.H){\overline \beta}][\ol{\alpha := \beta}] & = &
\lambda
d.\,\mathit{in}_{{T}^X_{H,\rho}} \;\;\mbox{ where } X \mbox{ is } \set \mbox{ when } \\ 
\hspace*{0.79in}(\mu \phi.\lambda {\overline \alpha}.H){\overline
  \beta} \rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in}  
\mathsf{D} = \set \mbox{ and not present when }
\mathsf{D} = \rel\\  
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash
  \fold^F_H : \Nat^\emptyset\;(\Nat^{\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F) & = &  
\lambda d\,\eta. \,\mathit{fold}_{T^X_{H,\rho}} \, \eta\\ 
\hspace*{0.79in}(\Nat^{{\ol{\beta}} }\,(\mu
  \phi.\lambda \overline \alpha.H)\overline \beta\;F)
\rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in} \mbox{ where } X \mbox{ is as above
}\vspace*{-0.1in} 
\end{array}\]
\caption{Term semantics}\label{fig:term-sem} 
\end{minipage}}\vspace*{-0.05in}
\end{figure*}

@@

\[
\begin{tikzcd}[row sep = huge]
\set^k
\ar[rr, "{F}"{name=Fa}, ""{name=F, below}]
\ar[rd, "{\ol K}"']
&& \set \\
& \set^h
\ar[Rightarrow, bend right = 10, from=F, "{\eta}" description]
\ar[ur, bend left, "{\mathit{Lan}_K\,F}"{name=Lan, description}]
\ar[ur, bend right, "{G}"'{name=L, right}, ""'{name=Lr, right}]
\ar[Rightarrow, bend left = 90, from=Fa, to=Lr, "{\gamma}"]
\ar[Rightarrow, dashed, from=Lan, to=L, "{\mu}"]
\end{tikzcd}
\]

@@

A third representation of left Kan extensions in locally presentable
categories is given in terms of colimits. Writing ${\mathcal C}_0$ for
the full subcategory of finitely presentable objects in the locally
presentable category $\mathcal C$, the left Kan extension can be
expressed as
\[
(\mathit{Lan}_{\ol K} F)\ol A \;=\; \colim{\ol{X : \,{\mathcal
        C}_0},\, \ol{f : K{\ol X} \to A}}{F \ol X}\]
Thus, if $F : \set^k \to \set$, $\ol{K} :
  \set^k \to \set^h$, $\ol{A} : \set^h$, and $\set_0$ is the full
  subcategory of finitely presentable objects in $\set$, --- i.e., is
  the category finite sets --- we  have that
\begin{equation}\label{eq:colim-form}
(\mathit{Lan}_K F) \ol{A} = \colim{\ol{S} : \set_0^k, \,\ol{f :
      K\ol{S} \to  A}}{F \ol{S}}
\end{equation}
If $\ol{S} : \set_0^k$ and $\ol{f : K\ol{S} \to A}$, let $j_{\ol{S},
  \ol{f}} : F \ol{S} \to (\mathit{Lan}_KF) \ol{A}$ be the morphism
indexed by $\ol{S}$ and $\ol{f}$ mapping the cocone into the colimit
in Equation~\ref{eq:colim-form}.
%Then for any $h$-tuple of functions
%$\ol{g : A \to B}$, the functorial action $(\mathit{Lan}_K F) \ol{g}$
%is the unique function from $(\mathit{Lan}_K F) \ol{A}$ to
%$(\mathit{Lan}_K F) \ol{B}$ such that, for all $\ol{S} : \set^k_0$ and
%$\ol{f : K\ol{S} \to A}$,\,
%\begin{equation}\label{eq:cocone-def}
%(\mathit{Lan}_KF) \ol{g} \circ j_{\ol{S},\, \ol{f}} = j_{\ol{S},\,
%    \ol{g\, \circ f}}
%  \end{equation}
%holds, where
For any $h$-tuple of functions
$\ol{g : A \to B}$, the functorial action $(\mathit{Lan}_K F) \ol{g}$
is the unique function from $(\mathit{Lan}_K F) \ol{A}$ to
$(\mathit{Lan}_K F) \ol{B}$ such that, for all $\ol{S} : \set^k_0$ and
$\ol{f : K\ol{S} \to A}$,\,
\begin{equation}\label{eq:cocone-def}
(\mathit{Lan}_KF) \ol{g} \circ j_{\ol{S},\, \ol{f}}
= j'_{\ol{S},\, \ol{g\, \circ f}}
: F \ol{S} \to (\mathit{Lan}_KF) \ol{B}
\end{equation}
holds, where $j$ is the morphism mapping the cocone into
$(\mathit{Lan}_K F) \ol{A} = \colim{\ol{S} : \set_0^k, \,\ol{f :
    K\ol{S} \to A}}{F \ol{S}}$ and $j'$ is the morphism mapping the
cocone into $(\mathit{Lan}_K F) \ol{B} = \colim{\ol{S} : \set_0^k,
  \,\ol{f : K\ol{S} \to B}}{F \ol{S}}$.  Moreover, if $\alpha : F \to
F'$ is a natural transformation, then $\mathit{Lan}_K \alpha :
(\mathit{Lan}_K F) \to (\mathit{Lan}_K F')$ is defined to be the
induced natural transformation whose component $(\mathit{Lan}_K
\alpha) \ol{A} : (\mathit{Lan}_K F) \ol{A} \to (\mathit{Lan}_K F')
\ol{A}$ is the unique function such that $(\mathit{Lan}_K \alpha)
\ol{A} \,\circ\, j_{\ol{S}, \ol{f}} = j'_{\ol{S}, \ol{f}} \,\circ\,
\alpha_{\ol{S}}$, where $j$ is the cocone into $(\mathit{Lan}_K F)
\ol{A}$ and $j'$ is the cocone into $(\mathit{Lan}_K F') \ol{A}$. We
can similarly represent left Kan extensions of functors over $\rel$ in
terms of colimits. In that setting we will denote the morphism mapping
the cocone into the colimit $\iota$. We will make good use of both of
these representations in Section~\ref{sec:term-sem} below.

@@

Intuitively, $\Lan^{\ol{\alpha^0}}_K F$ is a syntactic representation
of the left Kan extension of the functor in the variables in
$\ol\alpha$ denoted by $F$ along the functor in the variables in
$\ol\alpha$ denoted by $K$. Using $\Lan$-types, we can therefore
represent the GADT $\mathtt{Seq}$ in our calculus as 
 \begin{align*}
\mathtt{Seq}\,\alpha &\df \left( \mu \phi.\lambda \beta.
\beta + \left(\Lan^{\gamma_1,\gamma_2}_{\gamma_1 \times \gamma_2}
(\phi \gamma_1 \times \phi \gamma_2) \right) \beta \right. 
\left. + \left(\Lan^{\gamma}_{\mathtt{Nat} \to \gamma} (\mathtt{Nat} \to \phi
\gamma) \right) \beta \right) \alpha
\end{align*}
\noindent
where $\mathtt{Nat} = \mu \alpha. 1 + \alpha$ and $\mathtt{Nat}
\to C$ abbreviates $(\Lan^\alpha_{\mathtt{Nat}} 1)\,C$.  This
representation of $\mathit{Seq}$ corresponds exactly to the rewriting
in Section~\ref{sec:lke}.  As explained in Section~IV.\,D
of~\cite{jp19}, the more general construct $\Lan^{\ol{\alpha^0}}_{\ol
  K} F$ allowing extensions along vectors of functors as depicted
above makes it possible to represent GADTs with two or more type
arguments that depend on one another. Such GADTs cannot be represented
using just unary $\Lan$-types, i.e, $\Lan$-types of the form
$(\Lan^{\ol{\alpha^0}}_K F)\ol A$.

>>>>

We can prove by simultaneous induction that our interpretations of
types interact well with demotion of functorial variables to
non-functorial ones, along with other useful identities. Indeed, if
$\rho, \rho' : \setenv$, \,$f : \rho \to \rho'$, \,$\rho \phi = \rho
\psi = \rho' \phi = \rho' \psi$, \, $f \phi = f \psi = \id_{\rho
  \phi}$,\, $\Gamma; \Phi, \phi^k \vdash F$,\,
$\Gamma;\Phi,\ol{\alpha} \vdash G$,\, $\Gamma;\Phi,\alpha_1...\alpha_k
\vdash H$, and $\ol{\Gamma;\Phi \vdash K}$, then
\begin{gather}
\label{thm:demotion-objects}
\setsem{\Gamma; \Phi, \phi \vdash F} \rho = \setsem{\Gamma, \psi; \Phi
  \vdash F[\phi :== \psi] } \rho\\
\label{thm:demotion-morphisms}
\setsem{\Gamma; \Phi, \phi \vdash F} f = \setsem{\Gamma, \psi; \Phi
  \vdash F[\phi :== \psi]} f\\
\label{eq:subs-var}
\setsem{\Gamma;\Phi \vdash G[\ol{\alpha := K}]}\rho =
\setsem{\Gamma;\Phi,\ol{\alpha} \vdash G}\rho[\ol{\alpha := 
\setsem{\Gamma;\Phi \vdash K}\rho}]\\
\label{eq:subs-var-morph}
\setsem{\Gamma;\Phi \vdash G[\ol{\alpha := K}]}f =
\setsem{\Gamma;\Phi,\ol{\alpha} \vdash G}f[\ol{\alpha :=
\setsem{\Gamma;\Phi \vdash K}f}]\\
\label{eq:subs-const}
\setsem{\Gamma; \Phi \vdash F[\phi := H]}\rho
= \setsem{\Gamma; \Phi, \phi \vdash F}\rho
[\phi := \lambda \ol{A}.\, \setsem{\Gamma;\Phi,\overline{\alpha}\vdash
    H}\rho[\overline{\alpha := A}]] \\ 
\label{eq:subs-const-morph}
\setsem{\Gamma; \Phi \vdash F[\phi := H]}f
= \setsem{\Gamma; \Phi, \phi \vdash F}f
[\phi := \lambda \ol{A}.\,\setsem{\Gamma;\Phi,\overline{\alpha}\vdash
    H}f[\overline{\alpha := \id_A}]] 
\end{gather}
Identities analogous to (\ref{thm:demotion-objects}) through
(\ref{eq:subs-const-morph}) hold for relational interpretations as well.

@@

\begin{figure*}
\hspace*{-3.1in}
\resizebox{0.46\linewidth}{!}{
\begin{minipage}[t]{0.5\textwidth}
\[\begin{array}{lll}
\dsem{\Gamma;\Phi \,|\, \Delta,x :F \vdash x : F} \rho& = &
\pi_{|\Delta|+1}\\
\dsem{\Gamma;\emptyset \,|\, \Delta \vdash L_{\overline \alpha} x.t :
  \Nat^{\overline 
    \alpha} \,F \,G}\rho & = &  \curry (\dsem{\Gamma;\overline \alpha
  \,|\, \Delta, x : F \vdash t: G}\rho[\overline{\alpha := \_}])\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t_{\overline K} s:
  G [\overline{\alpha := K}]}\rho & = & \eval \circ \langle
  \lambda d.\,(\dsem{\Gamma;\emptyset \,|\, \Delta \vdash t :
  \Nat^{\overline{\alpha}} \,F \,G}\rho\; d)_{\overline{\dsem{\Gamma;\Phi
      \vdash K}\rho}},\\ 
 & & \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\,
    \Delta \vdash s: F [\overline{\alpha := K}]}\rho \rangle\\ 
& & \\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \bot_F t : F} \rho& = &
!^0_{\dsem{\Gamma;\Phi \vdash F}\rho} \circ
  \dsem{\Gamma;\Phi~|~\Delta \vdash t : \zerot}\rho, \mbox{ where } \\
 & & \hspace*{0.1in} !^0_{\dsem{\Gamma;\Phi \vdash F}\rho}
\mbox{ is the unique morphism from } 0\\
 & & \hspace*{0.1in} \mbox{ to } \dsem{\Gamma;\Phi \vdash F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \top : \onet}\rho & = &
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1, \mbox{ where }
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1\\ 
& & \hspace*{0.1in} \mbox{ is the unique morphism from }
\dsem{\Gamma;\Phi\vdash \Delta}\rho \mbox{ to } 1\\ 
\dsem{\Gamma;\Phi \,|\, \Delta \vdash (s,t) : F \times G} \rho& = &
\dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F} \rho\times
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G} \rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_1 t : F} \rho& = &
\pi_1 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times G}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_2 t : G}\rho & = &
\pi_2 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times
  G} \rho\\
\dsem{\Gamma;\Phi~|~\Delta \vdash \cse{t}{x \mapsto l}{y \mapsto r} :
  K}\rho & = & \eval \circ \langle \curry \,[\dsem{\Gamma;\Phi
    \,|\, \Delta, x : F \vdash l : K}\rho,\\
   & & \hspace*{0.79in} \dsem{\Gamma;\Phi \,|\, \Delta, y
    : G \vdash r : K}\rho],\\
   & &  \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\, \Delta \vdash t :
  F + G} \rho\rangle\\   
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inl \,s: F + G} \rho& = &
\inl \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inr \,t: F + G}\rho & = & 
\inr \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G}\rho\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \map^{\ol{F},\ol{G}}_H
  : \Nat^\emptyset (\Nat^{\ol\beta,\ol\gamma} F\,G)
& = & \lambda d\, \ol\eta\,\ol{C}.\,
\dsem{\Gamma; \ol{\phi},\ol{\gamma}\vdash H}\id_{\rho[\ol{\gamma:=
      C}]}[\ol{\phi := \lambda \ol B. \eta_{\ol B\, \ol C}}]\\ 
\hspace*{0.5in}
  (\Nat^{\ol{\gamma}}\,H[\ol{\phi :=_{\ol{\beta}} F}]\,H[\ol{\phi
      :=_{\ol{\beta}} G}]) \rrbracket^{\mathsf D} \rho & &\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
Nat^{\ol{\beta}} \, H[\phi := (\mu \phi.\lambda {\overline
    \alpha}.H){\overline \beta}][\ol{\alpha := \beta}] & = &
\lambda
d.\,\mathit{in}_{{T}^X_{H,\rho}} \;\;\mbox{ where } X \mbox{ is } \set \mbox{ when } \\ 
\hspace*{0.79in}(\mu \phi.\lambda {\overline \alpha}.H){\overline
  \beta} \rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in}  
\mathsf{D} = \set \mbox{ and not present when }
\mathsf{D} = \rel\\  
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash
  \fold^F_H : \Nat^\emptyset\;(\Nat^{\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F) & = &  
\lambda d. \,\mathit{fold}_{T^X_{H,\rho}} \\ 
\hspace*{0.79in}(\Nat^{{\ol{\beta}} }\,(\mu
  \phi.\lambda \overline \alpha.H)\overline \beta\;F)
\rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in} \mbox{ where } X \mbox{ is as above
}\vspace*{-0.1in} 
\end{array}\]
\caption{Term semantics}\label{fig:term-sem} 
\end{minipage}}\vspace*{-0.05in}
\end{figure*}

@@

\[
\begin{tikzcd}[row sep = huge]
\set^k
\ar[rr, "{F}"{name=Fa}, ""{name=F, below}]
\ar[rd, "{\ol K}"']
&& \set \\
& \set^h
\ar[Rightarrow, bend right = 10, from=F, "{\eta}" description]
\ar[ur, bend left, "{\mathit{Lan}_{\ol K}\,F}"{name=Lan, description}]
\ar[ur, bend right, "{G}"'{name=L, right}, ""'{name=Lr, right}]
\ar[Rightarrow, bend left = 90, from=Fa, to=Lr, "{\gamma}"]
\ar[Rightarrow, dashed, from=Lan, to=L, "{\mu}"]
\end{tikzcd}
\]

@@

A third representation of left Kan extensions in locally presentable
categories is given in terms of colimits. Writing ${\mathcal C}_0$ for
the full subcategory of finitely presentable objects in the locally
presentable category $\mathcal C$, the left Kan extension can be
expressed as
\[
(\mathit{Lan}_{\ol K} F)\ol A \;=\; \colim{\ol{X : \,{\mathcal
        C}_0},\, \ol{f : K{\ol X} \to A}}{F \ol X}\]
Thus, if $F : \set^k \to \set$, $\ol{K} :
  \set^k \to \set^h$, $\ol{A} : \set^h$, and $\set_0$ is the full
  subcategory of finitely presentable objects in $\set$, --- i.e., is
  the category finite sets --- we  have that
\begin{equation}\label{eq:colim-form}
(\mathit{Lan}_{\ol K} F) \ol{A} = \colim{\ol{S} : \set_0^k, \,\ol{f :
      K\ol{S} \to  A}}{F \ol{S}}
\end{equation}
If $\ol{S} : \set_0^k$ and $\ol{f : K\ol{S} \to A}$, let $j_{\ol{S},
  \ol{f}} : F \ol{S} \to (\mathit{Lan}_{\ol K} F) \ol{A}$ be the morphism
indexed by $\ol{S}$ and $\ol{f}$ mapping the cocone into the colimit
in Equation~\ref{eq:colim-form}.
%Then for any $h$-tuple of functions
%$\ol{g : A \to B}$, the functorial action $(\mathit{Lan}_K F) \ol{g}$
%is the unique function from $(\mathit{Lan}_K F) \ol{A}$ to
%$(\mathit{Lan}_K F) \ol{B}$ such that, for all $\ol{S} : \set^k_0$ and
%$\ol{f : K\ol{S} \to A}$,\,
%\begin{equation}\label{eq:cocone-def}
%(\mathit{Lan}_KF) \ol{g} \circ j_{\ol{S},\, \ol{f}} = j_{\ol{S},\,
%    \ol{g\, \circ f}}
%  \end{equation}
%holds, where
For any $h$-tuple of functions
$\ol{g : A \to B}$, the functorial action $(\mathit{Lan}_{\ol K
  }F) \ol{g}$
is the unique function from $(\mathit{Lan}_{\ol K} F) \ol{A}$ to
$(\mathit{Lan}_{\ol K} F) \ol{B}$ such that, for all $\ol{S} : \set^k_0$ and
$\ol{f : K\ol{S} \to A}$,\,
\begin{equation}\label{eq:cocone-def}
(\mathit{Lan}_{\ol K}F) \ol{g} \circ j_{\ol{S},\, \ol{f}}
= j'_{\ol{S},\, \ol{g\, \circ f}}
: F \ol{S} \to (\mathit{Lan}_{\ol K}F) \ol{B}
\end{equation}
holds, where $j$ is the morphism mapping the cocone into
$(\mathit{Lan}_{\ol K} F) \ol{A} = \colim{\ol{S} : \set_0^k, \,\ol{f :
    K\ol{S} \to A}}{F \ol{S}}$ and $j'$ is the morphism mapping the
cocone into $(\mathit{Lan}_{\ol K} F) \ol{B} = \colim{\ol{S} : \set_0^k,
  \,\ol{f : K\ol{S} \to B}}{F \ol{S}}$.  Moreover, if $\alpha : F \to
F'$ is a natural transformation, then $\mathit{Lan}_{\ol K} \alpha :
(\mathit{Lan}_{\ol K} F) \to (\mathit{Lan}_{\ol K} F')$ is defined to be the
induced natural transformation whose component $(\mathit{Lan}_{\ol K}
\alpha) \ol{A} : (\mathit{Lan}_{\ol K} F) \ol{A} \to (\mathit{Lan}_{\ol K} F')
\ol{A}$ is the unique function such that $(\mathit{Lan}_{\ol K} \alpha)
\ol{A} \,\circ\, j_{\ol{S}, \ol{f}} = j'_{\ol{S}, \ol{f}} \,\circ\,
\alpha_{\ol{S}}$, where $j$ is the cocone into $(\mathit{Lan}_{\ol K} F)
\ol{A}$ and $j'$ is the cocone into $(\mathit{Lan}_{\ol K} F') \ol{A}$. We
can similarly represent left Kan extensions of functors over $\rel$ in
terms of colimits. In that setting we will denote the morphism mapping
the cocone into the colimit $\iota$. We will make good use of both of
these representations in Section~\ref{sec:term-sem} below.

@@

Intuitively, $\Lan^{\ol{\alpha^0}}_{\ol K} F$ is a syntactic representation
of the left Kan extension of the functor in the variables in
$\ol\alpha$ denoted by $F$ along the functor in the variables in
$\ol\alpha$ denoted by $\ol K$. Using $\Lan$-types, we can therefore
represent the GADT $\mathtt{Seq}$ in our calculus as 
 \begin{align*}
\mathtt{Seq}\,\alpha &\df \left( \mu \phi.\lambda \beta.
\beta + \left(\Lan^{\gamma_1,\gamma_2}_{\gamma_1 \times \gamma_2}
(\phi \gamma_1 \times \phi \gamma_2) \right) \beta \right. 
\left. + \left(\Lan^{\gamma}_{\mathtt{Nat} \to \gamma} (\mathtt{Nat} \to \phi
\gamma) \right) \beta \right) \alpha
\end{align*}
\noindent
where $\mathtt{Nat} = \mu \alpha. 1 + \alpha$ and $\mathtt{Nat}
\to C$ abbreviates $(\Lan^\alpha_{\mathtt{Nat}} 1)\,C$.  This
representation of $\mathit{Seq}$ corresponds exactly to the rewriting
in Section~\ref{sec:lke}.  As explained in Section~IV.\,D
of~\cite{jp19}, the more general construct $\Lan^{\ol{\alpha^0}}_{\ol
  K} F$ allowing extensions along vectors of functors as depicted
above makes it possible to represent GADTs with two or more type
arguments that depend on one another. Such GADTs cannot be represented
using just unary $\Lan$-types, i.e, $\Lan$-types of the form
$(\Lan^{\ol{\alpha^0}}_{\ol K} F)\ol A$.
